test plot resize

### 2018-05-11 inst/documents/paper3/pap3_rot_analyses.Rmd
pap3

# I could maybe make the data longer and then facet by input
# think I would filter out, n_insecticides, start_freqs, male_expo_prop & rot_interval that don't show much rship
#TODO improve labels for these
inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos')
# something like this example from SO https://stackoverflow.com/a/48940750/1718356
#data_long <- tidyr::gather(data, key = type_col, value = categories, -col4)
#ggplot(data_long, aes(x = categories, fill = col4)) +
#  geom_bar() + 
#  facet_wrap(~ type_col, scales = "free_x")

>>>>

# I could maybe make the data longer and then facet by input
# think I would filter out, n_insecticides, start_freqs, male_expo_prop & rot_interval that don't show much rship
#TODO improve labels for these
inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos','start_freqs','rot_interval')
# something like this example from SO https://stackoverflow.com/a/48940750/1718356
#data_long <- tidyr::gather(data, key = type_col, value = categories, -col4)
#ggplot(data_long, aes(x = categories, fill = col4)) +
#  geom_bar() + 
#  facet_wrap(~ type_col, scales = "free_x")

### 2018-05-11 inst/documents/paper3/paper3_rotations.Rmd
pap3

Within the model inputs effect fitness as shown in Figure 1, differences in fitness lead to the changes in allele frequencies over time.

@@

@@

*todo change this bit, we could actually use time-to-resistance for all strategies ?*
*make it clearer exactly what we do*
It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative. For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5. This allowed comparison between strategies using a variable number of insecticides. Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used. Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.

@@

*show violin plots here ?*

@@

## TODO
*add to methods description of how fitness determined in model, move & modify fig 1*
*try to come up with brief name for model output to make it easier to refer to it later*
*add caveat about polygenic resistance*
*The caption to each graph needs to have the list of parameter values used to generate them*
*work on Ians fig Y to help explanation*

@@

```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia. The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence in which the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence reach the endpoint, with all insecticides having a resistance frequency above 0.5, at the same time. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE)
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)

@@

```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation lasts longer than the sequence. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step but in contrast to the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines  while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.05 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4. Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence last the same time. In both cases the resistance frequencies in the treated areas (red lines) increase when an insecticide is in use. Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas. When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1."}

linputs$n_insecticides <- 3
linputs$coverage <- 0.5
linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

ggrot <- rot_plot_resistance(dfrot, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=TRUE, logy=TRUE, plot=FALSE)
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)

@@

```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 5. Success of model runs under different scenarios and strategies. Model runs classed as a success if there are still insecticides for which resistance is below the threshold of 0.5 after 500 generations (circa 50 years for Anopheles)."}

#remove seq maybe have a check that are very few
# count(df_cost, success)
# 1         neither  5948
# 2             rot   209
# 3         rot_seq  3841
# 4             seq     2

# add a column for classification of whether, both, rotation, or neither succeeded
df_all <- mutate(df_all, success = ifelse(gens_seq==max_gen & gens_rot==max_gen,'both',
                                               ifelse(gens_rot==max_gen,'rotation only',
                                               ifelse(gens_seq==max_gen,'sequence only', 'neither'  
                                                      ))))
#count(df_all, success)
# 1          both 12413
# 2       neither 18196
# 3 rotation only   382
# 4 sequence only     9

@@

# plot with facets for dom_sel & dom_cos
ggplot(dfgrid, aes(x=freq, y=fitness_value, col=fitness_component)) +
  geom_line() +
  theme_minimal() +
  xlab('allele frequency') +
  facet_grid(dom_cos~dom_sel, labeller = "label_both", as.table = FALSE) #+
  # geom_text(data=abcd,
  #   aes(x = -Inf, y = +Inf, label=label),
  #   #inherit.aes=FALSE,
  #   colour='black',
  #   hjust = 0,
  #   vjust = 0.9 )

>>>>

*todo add to this description of how the model works and refer to SI where we can show the algebra*    
Within the model, the inputs described in Table 1 determine the fitness in each generation as shown in Figure 1, differences in fitness lead to the changes in allele frequencies in the following generation.

@@

*Ian : at this stage it would be good to revisit these model details before we continue further.*    
*It is fairly easy for me to rerun all the analyses if we change our mind on these.*

#### Optional model details    
Two optional model details were included to aid giving a more meaningful comparison of strategies. Firstly allele frequencies in a run were not allowed to go below the starting frequencies randomly chosen for that run. This aims to represent a situation where in the absence of insecticide use allele frequencies are maintained in balance between mutation and selection. Without this option enabled it was found that

Secoundly a constraint was added to stop rapid switching between insecticides every few generations which would be impractical in the field. Certain input combinations create this behaviour in the absence of a constraint. For example when two insecticides are close to the resistance frequency threshold of 0.5 it is possible for them to switch between each other every few generations as frequencies go above the threshold when in use and decline below it when not in use.  
*todo finish this part*

@@

*todo change this bit, think more about which outputs to use*
It is necessary to choose a measure to quantify the performance of an insecticide-use strategy and enable comparison with an alternative. For this we summed the number of generations when an insecticide was deployed and the resistance allele frequency for that insecticide was below 0.5. This allowed comparison between strategies using a variable number of insecticides. Previously time-to-resistance, namely the number of generations it takes to reach a defined resistance threshold (usually a resistance allele frequency of 0.5), was used. Using time-to-resistance worked when considering just two insecticides, but when the number of insecticides is a variable it is not sufficient.

@@

*refer to violin plots here ?*

@@

## TODO

1. *add to methods description of how fitness determined in model, move & modify fig 1*
1. *add caveat about polygenic resistance*

@@

```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia. The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence in which the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence reach the endpoint, with all insecticides having a resistance frequency above 0.5, at the same time. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.7, exposure=0.7, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  150
linputs$expo_hi <-  0.7
linputs$eff <-      0.7
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='A. Rotation')
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='B. Sequence')
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)

@@

```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation lasts longer than the sequence. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step but in contrast to the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines  while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.7, exposure=0.7, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1."}

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.05 
linputs$max_gen <-  150
linputs$expo_hi <-  0.7
linputs$eff <-      0.7
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='A. Rotation')
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=TRUE, plot=FALSE, title='B. Sequence')
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)

```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4. Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence last the same time. In both cases the resistance frequencies in the treated areas (red lines) increase when an insecticide is in use. Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas. When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.7, exposure=0.7, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 0.5
linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  150
linputs$expo_hi <-  0.7
linputs$eff <-      0.7
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=4, fig.show='hold', fig.align='center', fig.cap="Figure 5. Success of model runs under different scenarios and strategies. Model runs classed as a success if there are still insecticides for which resistance is below the threshold of 0.5 after 500 generations (circa 50 years for Anopheles). With no costs or dispersal both strategies succeeded in around 5% of runs and neither did in 95%.  When costs alone were included the proportion of runs in which both, neither and rotation succeeded was 38%, 60% and 2%. When dispersal alone was included both strategies succeeded in 20% or runs and neither did in 80%, in only 0.1% of runs did rotation alone succeed."}

# add a column for classification of whether, both, rotation, or neither succeeded
df_all <- mutate(df_all, success = ifelse(gens_seq==max_gen & gens_rot==max_gen,'both',
                                               ifelse(gens_rot==max_gen,'rotation only',
                                               ifelse(gens_seq==max_gen,'sequence only', 'neither'  
                                                      ))))
#count(df_all, scenario, success)
# 1 no cost or dispersal          both    47
#  2 no cost or dispersal       neither   952
#  3 no cost or dispersal sequence only     1
#  4                costs          both  3841
#  5                costs       neither  5948
#  6                costs rotation only   209
#  7                costs sequence only     2
#  8            dispersal          both  1938
#  9            dispersal       neither  8046
# 10            dispersal rotation only    12
# 11            dispersal sequence only     4
# 12  costs and dispersal          both  6587
# 13  costs and dispersal       neither  3250
# 14  costs and dispersal rotation only   161
# 15  costs and dispersal sequence only     2

@@

```{r Fig5xtra, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.cap="Figure 5xtra. How model runs classed as success under the costs only scenario for the different strategies differ in the values of inputs. Violin plots showing how the density of input values changes according to the classification of success. Runs in which both strategies were successful were more likely to have higher costs and lower effectiveness, exposure and resistance restoration. Runs in which neither strategy was successful were the reverse. Runs for which rotations only were successful tended to have higher dominance of cost, lower dominance of resistance and lower rotation intervals. Remember that as shown in the previous figure the proportion of runs for which rotation alone succeeded was lower than for the other options."}

# I could maybe make the data longer and then facet by input
# think I would filter out, n_insecticides, start_freqs, male_expo_prop & rot_interval that don't show much rship
#TODO improve labels for these
inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos','start_freqs','rot_interval','male_expo_prop')
# something like this example from SO https://stackoverflow.com/a/48940750/1718356
#data_long <- tidyr::gather(data, key = type_col, value = categories, -col4)
#ggplot(data_long, aes(x = categories, fill = col4)) +
#  geom_bar() + 
#  facet_wrap(~ type_col, scales = "free_x")

#filter cost from all because the success calc was done on that
df_cost_long <- filter(df_all,scenario=='costs')
#first just select the columns I want
df_cost_long <- select(df_cost_long,c(inputs,'success'))
#gather data to be longer
df_cost_long <- tidyr::gather(df_cost_long, key=input_name, value=input_value, -success)

#there were just an odd couple of results where seq better, remove them here or earlier
df_cost_long <- filter(df_cost_long, success != 'sequence only')

ggplot(df_cost_long, aes_string(x='success',y='input_value', color='success')) + 
        #ylim(0,450) +
        #coord_cartesian( ylim=c(0, 350)) +
        geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = 1 ) +
        theme_minimal() +
        #this is an example of how you might facet
        facet_wrap(~input_name, scales = "free_y") +  
        theme(panel.grid.minor.y=element_blank()) #,
        #           axis.text.x=element_blank())
        #geom_jitter(shape='.',height=0,width=0.2) )

```{r Fig5xtra_cost_dispersal, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.cap="Figure 5xtra cost and dispersal. Probably don't show in the paper. Similar to previous one. Just adds coverage and migration."}

#TODO improve labels for these

#coverage and migration added for costs and dispersal
inputs <- c('cost', 'expo_hi', 'eff', 'rr', 'dom_sel', 'dom_cos','start_freqs','rot_interval','male_expo_prop','coverage','migration')

#filter cost from all because the success calc was done on that
df_cost_long <- filter(df_all,scenario=='costs and dispersal')
#first just select the columns I want
df_cost_long <- select(df_cost_long,c(inputs,'success'))
#gather data to be longer
df_cost_long <- tidyr::gather(df_cost_long, key=input_name, value=input_value, -success)

#there were just an odd couple of results where seq better, remove them here or earlier
df_cost_long <- filter(df_cost_long, success != 'sequence only')

ggplot(df_cost_long, aes_string(x='success',y='input_value', color='success')) + 
        #ylim(0,450) +
        #coord_cartesian( ylim=c(0, 350)) +
        geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = 1 ) +
        theme_minimal() +
        #this is an example of how you might facet
        facet_wrap(~input_name, scales = "free_y") +  
        theme(panel.grid.minor.y=element_blank()) #,
        #           axis.text.x=element_blank())
        #geom_jitter(shape='.',height=0,width=0.2) )

@@

# plot with facets for dom_sel & dom_cos
ggplot(dfgrid, aes(x=freq, y=fitness_value, col=fitness_component)) +
  geom_line() +
  theme_minimal() +
  xlab('allele frequency') +
  scale_x_continuous(breaks=seq(0,1,0.2)) +
  facet_grid(dom_cos~dom_sel, labeller = "label_both", as.table = FALSE) #+
  # geom_text(data=abcd,
  #   aes(x = -Inf, y = +Inf, label=label),
  #   #inherit.aes=FALSE,
  #   colour='black',
  #   hjust = 0,
  #   vjust = 0.9 )

### 2018-05-11 inst/documents/rotations_notes.Rmd
pap3

*10/5/18*

Testing setting width of plots for publication at 300 dpi into word doc.

@@

@@

>>>>

*10/5/18*
Testing setting width of plots for publication at 300 dpi into word doc.

@@

got Ians fig Y working in freq_fitness_simple() & added facetted grid plot to the paper.

*11/5/18*
- improved violin plots added to pap3
- improving legends

Finally grasping the nettle of insecticide switching.

I want to stop switching to an insecticide that was used < X generations ago.

To do that store an array which has a counter for each insecticide which records how long since it was last used. Then just add to that counter in each generation for each insecticide not in use.
Don't allow an insecticide to be used that is below a user-defined level.

in insecticide_check()
BE CAREFUL!

df_ins <- data.frame(last_used=rep(NA, n_insecticides))

@@

Before sending to Ian.

~ add to methods about min rwr interval
~ add about keeping resistance allele freq at starting level
~ add to fig Y legend Ians caveats
